---
title: "plot"
author: "Xinge Zhang"
date: "2023-02-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Ⅰ.Introduction

### Background

### Motivation & Use Case

## Ⅱ. Exploratory Analysis

### The Data: Understanding landcover data

#### dependent: pervious to impervious surface during the period 2014-201
#### independent:
    lccover change 能change
    1. `originallc` originallandcover
    
    2. `lcp` permeable/impermeable:
    < 6 - 0
    >= 6 - 1
    
    3. `lcchange` whether landcover has changed: 1
    
    4. `lc` 3*3的permeable rate:
    0-1（0.1，0.2， 0.3， 0.4， 0.5， 0.6， 0.7， 0.8）
    
    5. `popchange` per m^2
    
    6. `pctwhitechange`
    
    7. `unitchange` per m^2
    
    8. `medhhincchange` per m^2
    
    9. `road` road 1/0
    
    10. `water` water
    
    11. `canopy` tree canopy：
    tree canopy - 1
    tree canopy over structure -0.5
    tree canopy over other -0.5
    tree canopy over road - 0.5
    other - 0
    
    12. `perm` permeable:
    2 -shrub - 0.3
    4 - herb - 0.3
    5 - wetand - 0.4
    other -0
    
    13. `barren` barren:
    6 - barren - 1
    not barren - 0
    
    14. `struct` structure:
    7 - structure - 1
    not structure - 0
    
    15. `other` other:
    8 - other - 1
    not other - 0
    
    16. `terrain`dem
    
    17. `slope`slope
    
    18. `area` m^2 per block group
    
    19. `geoid`-- blockgroup id
    
    20. `soil type`-


模型encode：categorial：尤其是

possible features per 10*10 :

Road density
**distance to rivers**

impervious / pervious
landcover type
population density
population change per 4 year
median hh income
median hh income change per 4 year
percentage of white population
percentage of white population change per 4 year
percentage of changing from impervious to pervious/from pervious to impervious in 4 years
predicted landcover type:

possible features per block group :

percentage of impervious / pervious change:
population
population change per 4 year
median hh income
median hh income change per 4 year
percentage of white population
percentage of white population change per 4 year
percentage of changing from impervious to pervious/from pervious to impervious in 4 years
predicted landcover type composition:

possible feautres per county:

percentage of impervious / pervious change:
population
population change per 4 year
median hh income
median hh income change per 4 year
percentage of white population
percentage of white population change per 4 year
percentage of changing from impervious to pervious/from pervious to impervious in 4 years
predicted landcover type composition:
development speed defined by: speed of population change and landcover change speed
Agricultural land use, vegetation cover and water loss:


```{r}
glimpse(isle18_df)
```

```{r}

glimpse(isle14_df)
```
```{r fig.width=15}
# Define column names vector
col_names <-colnames(isle14_df)[5:8] 

# Set up empty list to store plots
plots_list <- list()

# Set up for loop
for (col in col_names) {
  
  # Create plot for current column
  p <- ggplot(isle14_df, aes(x = isle14_df[,col],fill = lcp, group= lcp)) +
    geom_density(alpha = .3) +
    xlab("Value") +
    ggtitle(col)
  
  # Add plot to list
  plots_list[[col]] <- p
}

# Display plots in a grid format
grid.arrange(grobs = plots_list, nrow = length(col_names))

```


```{r fig.width=15}
# Define column names vector
col_names <-colnames(isle14_df)[9:12] 

# Set up empty list to store plots
plots_list <- list()

# Set up for loop
for (col in col_names) {
  
  # Create plot for current column
  p <- ggplot(isle14_df, aes(x = isle14_df[,col],fill = lcp, group= lcp)) +
    geom_histogram(alpha = .3) +
    xlab("Value") +
    ggtitle(col)
  
  # Add plot to list
  plots_list[[col]] <- p
}

# Display plots in a grid format
grid.arrange(grobs = plots_list, ncol= length(col_names))

```




```{r}
ggplot(isle14_df, aes(x= pctwhitechange, fill = lcp, group= lcp)) +
   geom_density(alpha = .3)
```

